Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-21 ind_files/05_unification_discussion.tex
Minor rephrasing

If the `intuitive' version of It\^{o}'s formula for $\mathcal{M}_K(\mathcal{T})$ valued stochastic processes obtained by `taking the limit' $m \to \infty$ in the It\^{o}'s formula for $m$-dimensional stochastic processes holds, then the exact same steps carried out in Appendices \ref{App_density_to_freq} and \ref{App_stoch_var_eqns} will `go through' essentially unchanged for quantitative traits and yield the equations obtained by simply taking $m \to \infty$ in \eqref{nD_stochastic_RM}, \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} (with sums replaced by integrals and derivatives replaced by functional derivatives as appropriate) as the `fundamental theorems' for the evolution of quantitative traits. In their supplementary sections S1 and S2,~\cite{week_white_2021} have recently proposed exactly the It\^o formula we would need for $L^2(\mathbb{R}, m)$ valued processes in which the $\mathcal{A}^{\pm}(x|\phi)$ (in my notation as defined in chapter \ref{chap_infD_processes}) are Hilbert-Schmidt operators via a heuristic It\^o multiplication table, and have shown that their heuristics are equivalent to the rigorous infinite-dimensional stochastic calculus proposed by~\cite{da_prato_stochastic_2014} for more general Hilbert space valued processes\footnote{If this sentence sounds like abstract gibberish to you, don't worry too much about the details - The essence is that~\cite{week_white_2021} have proven the `correct' formula we need holds whenever $b(x|\nu), d(x|\nu)$ and $\mathcal{M}_K(\mathcal{T})$ together fulfill certain technical requirements. My heuristics here are only a `first step' and have focused on accessibility and actually formulating the problem in formal terms rather than on mathematical propriety, and thus make no attempt to verify whether and when these technical requirements are satisfied. The required properties will likely be fulfilled only for some `nice' subspace of $\mathcal{M}_K(\mathcal{T})$ and not the entire space, and future, more rigorous work should focus on figuring out which functions are included in this subspace.}. Furthermore, using these `spacetime white noise heuristics' together with a particular functional form for (weak) solutions to an SPDE describing populations with Gaussian trait distributions,~\cite{week_white_2021} have arrived at precisely the infinite-dimensional version of equations \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} obtained by `taking the limit' $m \to \infty$ in \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} for the special case in which the type-level quantity $f$ is simply the value of the quantitative trait under study from a completely different starting point(!). Equation \eqref{disc_functional_field_eqns} is the `functional Fokker-Planck view' of the stochastic processes for the evolutionary ecology of quantitative traits that~\cite{week_white_2021} study from the `SPDE view'. Indeed, the two approaches are exactly complementary, and my formulation in chapter \ref{chap_infD_processes} provides an alternate method of attack for the study of quantitative traits in finite fluctuating populations that may be more appropriate for some particular problems, while~\cite{week_white_2021}'s approach may be more appropriate for others.

>>>>

If the `intuitive' version of It\^{o}'s formula for $\mathcal{M}_K(\mathcal{T})$ valued stochastic processes obtained by informally `taking the limit' $m \to \infty$ in the It\^{o}'s formula for $m$-dimensional stochastic processes holds, then the exact same steps carried out in Appendices \ref{App_density_to_freq} and \ref{App_stoch_var_eqns} will `go through' essentially unchanged for quantitative traits and yield the equations obtained by simply informally taking $m \to \infty$ in \eqref{nD_stochastic_RM}, \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} as the `fundamental equations' for the evolution of quantitative traits. In their supplementary sections S1 and S2,~\cite{week_white_2021} have recently proposed exactly the It\^o formula we would need for $L^2(\mathbb{R}, m)$ valued processes in which the $\mathcal{A}^{\pm}(x|\phi)$ (in my notation as defined in chapter \ref{chap_infD_processes}) are Hilbert-Schmidt operators via a heuristic It\^o multiplication table, and have shown that their heuristics are equivalent to the rigorous infinite-dimensional stochastic calculus proposed by~\cite{da_prato_stochastic_2014} for more general Hilbert space valued processes\footnote{If this sentence sounds like abstract gibberish to you, don't worry too much about the details - The essence is that~\cite{week_white_2021} have proven the `correct' formula we need holds whenever $b(x|\nu), d(x|\nu)$ and $\mathcal{M}_K(\mathcal{T})$ together fulfill certain technical requirements. My heuristics here are only a `first step' and have focused on accessibility and actually formulating the problem in formal terms rather than on mathematical propriety, and thus make no attempt to verify whether and when these technical requirements are satisfied. The required properties will likely be fulfilled only for some `nice' subspace of $\mathcal{M}_K(\mathcal{T})$ and not the entire space, and future, more rigorous work should focus on figuring out which functions are included in this subspace.}. Furthermore, using these `spacetime white noise heuristics' together with a particular functional form for (weak) solutions to an SPDE describing populations with Gaussian trait distributions,~\cite{week_white_2021} have arrived at precisely the infinite-dimensional version of equations \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} obtained by `taking the limit' $m \to \infty$ in \eqref{nD_stochastic_Price} and \eqref{nD_stochastic_Price_variance} for the special case in which the type-level quantity $f$ is simply the value of the quantitative trait under study from a completely different starting point(!). Equation \eqref{disc_functional_field_eqns} is the `functional Fokker-Planck view' of the stochastic processes for the evolutionary ecology of quantitative traits that~\cite{week_white_2021} study from the `SPDE view'. Indeed, the two approaches are exactly complementary, and my formulation in chapter \ref{chap_infD_processes} provides an alternate method of attack for the study of quantitative traits in finite fluctuating populations that may be more appropriate for some particular problems, while~\cite{week_white_2021}'s approach may be more appropriate for others.
